Probabilistic Symbol Encoding for \\Convolutional Associative Memories

### 2017-01-20 papers/hrr-paper/paper.tex
Corrected Symbolic Scrypting Experiment, rearranged images as well.

We have used our system to demonstrate a mechanism for encoding neural programs and processes through a series of associated action, perception and actuator symbols. We call this mechanism for defining controllers using holographic memories "symbolic scripting".

@@

The basic idea is centered around symbolic representations of all three parts involved:
\begin{enumerate}
	\item \textbf{actions} are symbolic constants representing groups of control signals, such as "forward", "backward", "stop", "flex", etc.

@@

An example of this kind of symbolic script is given in figure \ref{fig:symbolic-scripting}. Here the memory encodes an object following behavior for left wheel.

		\begin{figure}[th!]
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/gaussianbar.png}
				\caption{Gaussian probability distribution for raw input value\\}
				\label{fig:gaussianbar}
			\end{subfigure}
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/gaussianbarperm.png}
				\caption{Probabilistic symbolic encoding suitable for HRR after permutation
				}
				\label{fig:gaussianbarperm}
			\end{subfigure}	
		\caption{Sensory input encoding of target object position in the visual field at two successive points in time. Left graph shows original encodings for two adjacent values and right graph shows permuted encodings. It can be seen that semantically close input values result in similar HRR encodings.}
		\label{fig:gaussian}
		\end{figure}

@@

The symbolic scripting experiment is set up using a single-sensor Braitenberg
vehicle (see figures \ref{fig:experiment:agent1} and
\ref{fig:experiment:agent2}).
The sensor returns a list of objects in its field of view with their
corresponding view space positions in $[0,1]$ (far left to far right).
% note: it would be better if we probe with all returned objects and check the
% result, but IIRC this did not work well. maybe figure it out?
All returned objects are then compared to the agent's target object.
If the target is found, the holographic controller is probed with its view space
position which, due to the similarity of scalar values, yields results even for
values that lie between the originally stored values and can consequently be
probed with the symbols $left\_wheel$ and $right\_wheel$ to extract the symbolic
actions $forwards$ or $backwards$.
If the target is not found or no action could be determined, the agent defaults
to probing the controller with the symbolic constant $no\_object$ which results
a clockwise rotation when probing for wheel actions (see figure
\ref{fig:experiment:similar2}).

In the left column of figure \ref{fig:experiment}, the agent approaches the blue
target (figure \ref{fig:experiment:agent1}). In the beginning (tick $-50$), the
agent rotates right (default action) because the target is not visible.
At tick $-30$, the object enters the agent's visual field and is consequently
centered until tick $-19$.
Then, with the sensory value of ${\sim}0.5$, the actions for both wheels switch
to $forwards$ (figure \ref{fig:experiment:similar1}) and the agent approaches
the target, constantly decreasing its distance
(figure \ref{fig:experiment:distance1}).

In the right column of figure \ref{fig:experiment}, the target is not currently
in the visual field of the agent.
This is due to reaching its old target at approximately tick $-13$ which leads
to the removal of all current objects and random placement of new objects,
including one target.
As the new object is placed outside of the agent's field of view, the default
action of probing $no\_object$ is performed constantly.
As described, this then results in the encoded default actions $backwards$ for
probing with $left\_wheel$ and $forwards$ for probing with $right\_wheel$---a
clockwise rotation (see figure \ref{fig:experiment:similar2}).

This experiment shows the effectiveness of our approach in a practical
application by encoding symbolic sensing, actuators and actions in a symbolic
controller.

@@

	\caption{Scene of Agent 1 with three randomly placed items, the blue

@@

	The agent does not have the target object in his field of view (green
	fan).}

@@

	\caption{Symbolic sensing with target in visual field results in scalar
	value representing target direction (${\sim}0.55$, almost straight

@@

	\caption{Symbolic sensing without target in visual field showing the
	representation for the symbolic constant $no\_object$.}

@@

	\caption{Target position over time in the visual field of Agent 1.
	At tick $-30$, it appears on the far right edge and then gradually

@@

	In this case, the old target was reached and the new target is not in
	the field of view yet.}

@@

	Here, the blue object in \subref{fig:experiment:agent1} is being

@@

	At tick $-12$ a new target was added (black bar).}

@@

	\caption{Similarities as described in \subref{fig:experiment:similar1}

@@

	the symbol $no\_object$ is performed (rotate right).}

@@

The left column (\protect\subref{fig:experiment:agent1},

@@

\protect\subref{fig:experiment:similar2}) shows an agent without a target in its

>>>>

We have used our system to create a mechanism for encoding neural programs and processes through a series of associated action, perception and actuator symbols. We call this mechanism for defining controllers using holographic memories "symbolic scripting".

@@

The basic idea is centred around symbolic representations of all three parts involved:
\begin{enumerate}
	\item \textbf{actions} are symbolic constants representing groups of control signals, such as "forward", "backward", "stop", "flex", etc.

@@

An example of this kind of symbolic script is given in figure \ref{fig:symbolic-scripting}. Here the memory encodes an object following behaviour for its left wheel.

		\begin{figure}[th!]
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/gaussianbar.png}
				\caption{Gaussian probability distribution for raw input value\\}
				\label{fig:gaussianbar}
			\end{subfigure}
			\begin{subfigure}{0.45\columnwidth}
				\center
				\includegraphics[width=1\columnwidth]{img/gaussianbarperm.png}
				\caption{Probabilistic symbolic encoding suitable for HRR after permutation
				}
				\label{fig:gaussianbarperm}
			\end{subfigure}	
		\caption{Sensory input encoding of a target objects position in the visual field at two successive points in time. Left graph shows original encodings for two adjacent values and right graph shows permuted encodings. It can be seen that semantically close input values result in similar HRR encodings.}
		\label{fig:gaussian}
		\end{figure}

@@

The symbolic scripting experiment is set up using a single-sensor Braitenberg
vehicle (see figures \ref{fig:experiment:agent1} and
\ref{fig:experiment:agent2}).
The sensor returns a list of objects in its field of view with their
corresponding view space positions in $[0,1]$ (far left to far right).
% note: it would be better if we probe with all returned objects and check the
% result, but IIRC this did not work well. maybe figure it out?
All returned objects are then compared to the agent's target object.
If the target is found, the holographic controller is probed with its view space
position which, due to the similarity of scalars, yields results even for
values that lie between the originally stored values and can consequently be
probed with the symbols $left\_wheel$ and $right\_wheel$ to extract the symbolic
actions $forwards$ or $backwards$.
If the target is not found or no action could be determined, the agent defaults
to probing the controller with the symbolic constant $no\_object$ which results in
a clockwise rotation when probing for wheel actions (see figure
\ref{fig:experiment:similar2}).

In the left column of figure \ref{fig:experiment}, the agent approaches the blue
target (figure \ref{fig:experiment:agent1}). In the beginning (tick $-50$), the
agent rotates right (default action) because the target is not visible.
At tick $-30$, the object enters the agent's visual field and continuously nears the fields center, which is reached at tick $-19$.
Then, with the sensory value of ${\sim}0.5$, the actions for both wheels switch
to $forwards$ (figure \ref{fig:experiment:similar1}) and the agent approaches
the target, constantly decreasing its distance
(figure \ref{fig:experiment:distance1}).

In the right column of figure \ref{fig:experiment}, the target is not yet
in the visual field of the agent.
This is the result of the agent previously reaching an old target (at approximately tick $-13$) which lead
to the removal of all current objects and random placement of new ones,
including a new target.
As the new target is placed outside of the agent's field of view, the default
action of probing $no\_object$ is performed. The
clockwise rotation can also be seen in figure \ref{fig:experiment:similar2}). Regardless of how many times the agent reaches its target it will continue to find and pursue new ones solely with the help of its symbolic controller.

This experiment shows the effectiveness of our approach in a practical
application by encoding symbolic sensing, actuators and actions in a
controller.

@@

	\caption{Scene of Agent 1 (red) with three randomly placed items, the blue

@@

	The agent does not have the target object in his field of view yet.}

@@

	\caption{Symbolic sensing with the target in the visual field, which results in a scalar
	value representing the target direction (${\sim}0.55$, almost straight

@@

	\caption{Symbolic sensing without the target in the visual field showing the
	representation for the symbolic constant $no\_object$.\\}

@@

	\caption{The targets position over time in the visual field of Agent 1.
	At tick $-30$, the target enters the far right edge of the field of view and then gradually

@@

	This is due to it having been reached, which prompts a new target creation, which is not in
	the field of view.}

@@

	Here, the blue object in sub-figure \subref{fig:experiment:agent1} is being

@@

	At tick $-12$ a new target was created (black bar).}

@@

	\caption{Similarities as described in sub-figure \subref{fig:experiment:similar1}

@@

	the symbol $no\_object$ is performed (rotate right).\\}

@@

The left column (sub-figures \protect\subref{fig:experiment:agent1},

@@

\protect\subref{fig:experiment:similar2}) shows an agent with currently no target in its
